Manuscript

### 2022-08-27 Manuscript.Rmd
new discussion approach HP

---
title: "Manuscript"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: 
  - Justin Pomeranz:
      email: jfpomeranz@gmail.com
      institute: cmu
      correspondence: true
  - name: James R. Junker
    institute: [glrc, lumcon]
  - name: Jeff S. Wesner
    institute: usd
institute:
  - cmu: Colorado Mesa University
  - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - lumcon: Louisiana Universities Marine Consortium, Chauvin, LA USA
  - usd: Dept. of Biology, University of South Dakota, Vermillion, SD, USA
output:
   pdf_document:
     toc: yes
     keep_tex: true
     pandoc_args:
     - '--lua-filter=scholarly-metadata.lua'
     - '--lua-filter=author-info-blocks.lua'
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   html_document: default
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
  - \usepackage{lineno}
  - \usepackage{amsmath}
  - \usepackage{indentfirst}
  - \linenumbers
  - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

For each sample of $M$ and each of the 1000 replicates, we estimated the exponent $\lambda$ using MLE methods modified from the the `sizeSpectra` package (Edwards et al. 2020). In addition, we use two common binning methods to estimate the OLS slope parameter in log-log space. For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes. The count in each bin was normalized by dividing by the bin width. This method has been used by (refs here, I think a lot of Woodward's group uses this...) Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn. The second method was similar to ELBn, but bins of Log~2~ widths are used. The count in each bin is normalized in the same way. This method has been used by (Pomeranz et al. 2018, I **think** McGarvey \~2018, Blanchard (log 2, ?normalized?))[@pomeranz2018; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).

@@

For each simulation replicate, we estimated how the size spectra parameters varied across the hypothetical gradient. Simple OLS regression were conducted in the form $\lambda_{estimate} ~ \beta_0 + \beta_1 * X$. The distribution of the relationship coefficient, $\beta_1$, were plotted compared to the known relationship. Likewise, the confidence interval of $\beta_1$ was assessed to determine the proportion of estimates which contained the true value of the known relationship.

@@

We re-analyze two datasets of benthic macroinvertebrate communities from stream habitats across two different gradients. In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage stress gradient. Details of the sample collection and processing can be found in Pomeranz et al. (2018)[@pomeranz2018]. Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope. Body mass was estimated using published length weight regressions.

The second data set was from the wadeable stream sites of the National Ecological Observatory Network (NEON; **data product XXX**)[@NEON_Inverts2022]. These sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska. Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat. All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions. This data has been analyzed previously using size spectra methods as described in Pomeranz et al. (2022)[@pomeranz2022]. Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([https://www.neonscience.org/data-collection/macroinvertebrates], or macroinvertebrate DPI pubs).

@@

All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient. The regressions using the MLE method found a significant relationship 100% of the time, while the ELBn and NAS method found significant relatoinships 95.8% and 96.9% of the time, respectively.

The confidence intervals for the relationship coefficients ($\beta_1$) for both the MLE and ELBn method contained the true value of the known relationship across the gradient $\ge$ 95% of the time, whereas the confidence interval for the NAS method only had the true value 83.1% of the time. Despite having similar accuracy, the width of the CIs for the ELBn method were more than 3 times that of the MLE method (ELBn_CI_ = 0.386 $\pm$ 0.20; MLE_CI_ = 0.121 $\pm$ 0.056). The CI's for the NAS were slightly smaller than the ELBn method, but still $>$ 2.5 times as wide as the MLE method (NAS_CI = 0.339 $\pm$ 0.172).

On average, the relationship across the gradient was under estimated: MLE -0.002 ($\pm$ 0.022); ELBn -0.039 ($\pm$ 0.062); and NA -0.081 ($\pm$ 0.061).

@@

All of the methods performed similarly when there was no relationship across the hypothetical gradient. The Type I error rate for MLE, ELBn and NAS were 5%, 5.7% and 4.6%, respectively. However, the confidence intervals for the binning methods were \~3 times as wide as the MLE method (mean CI widths: MLE = 0.116, ELBn = 0.372; NAS = 0.332).

Of the relationships which were estimated to be significant, the MLE, ELBn, and NAS method had a 58%, 49.1% and 54.3% probability of indicating a negative relationship, respectively.

@@

The performance of all methods declined when trying to detect variation in the $\lambda$ parameter between -1.9 to -2.1 across the environmental gradient. The mean $\beta_1$ coefficient estimates for all methods were closely associated with the known relationship. Once again the variation in the estimates using the MLE method were much smaller that the estimates using both of the binning methods (Figure 4).

![Mean difference between relationship estimate and known relationship coefficient for each method. Here, $\lambda$ varied from -1.9 to -2.1, with a kown relationaship of -0.1 across the gradient.](figures/PLB_sim_small_relationship_density.png)

Of the 1000 simulation replicates, the MLE only detected a significant relationship 90% of the time, whereas the ELBn and NAS methods detected significant relationships 19 and 23% of the time, respectively. Of the replicates which were significant, the confidence interval contained the true value of the relationship 94.6, 88.4, and 89.1% of the time for the MLE, ELBn, and NAS methods, respectively.

@@

Once again, the width of the confidence intervals for both of the binning methods were $\ge$ 2.8 times that of the MLE method (mean $\pm$ SD MLE CI width: 0.117 $\pm$ 0.051; ELBn CI width: 0.378 $\pm$ 0.169; NAS CI width: 0.332 $\pm$ 0.150).

On average, the relationship estimates across the gradient were similar to the know value (mean deviation: MLE = 0.00001, ELBn = -0.009, NAS = -0.012). However, when only looking at the relationships which were significant, this variation increased considerably(MLE = -0.002, ELBN = -0.052, NAS = 0.034).

@@

![D.](figures/neon_relationship.png)

@@

Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology. Previous work has investigated the accuracy and inherent biases associated with different estimation methods. However, how these innaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detecting variation in size spectra parameters across environmental gradients with confidence. Here, we sampled body sizes from known distributions and estimated their parameters across a hypothetical gradient in order to assess the relationship coefficients obtained.

Generally, there was large agreement in the direction and magnitude of average relationship coefficient estimates across methods. The estimate from the MLE method was always closer to the known relationship, and always had the smallest variation around the estimate.

All of the methods performed equally well when the $\lambda$ value did not vary across the hypothetical gradient, with a Type I error rate of \~5%, exactly as would be expected under the assumptions of frequentist statistics. This indicates that previous studies which have detected no difference in size spectra parameters across sites can be interpreted with confidence, regardless of the method used.

@@

Alarmingly, the performance of all methods declined dramatically when trying to detect smaller variation in $\lambda$ values. We simulated observations from distributions varying by an absolute value of 0.2 units. This range of variation is similar to that which has been reported in many seminal works on variation in size spectra parameters. When using the MLE method, significant relationships were only detected \~88% of the time, whereas both of the binning methods only detected significant trends \~20% of the time. Of the simulation replicates which were significant, nearly all of them were negative across all methods except for one single replicate using the NAS method which was positive (Figure. 5).

Despite the drop in performance with reduced variation in $\lambda$ values, when a significant relationship was observed, it was generally in the correct direction and of a similar magnitude. This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible, and the magnitude of the relative change is a reasonable estimate. Given that all of the data within a study is treated identically, the the over all chnage in size spectra parameters is likely reasonable. However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to compare the relative changes across different published studies which use different methods. Publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental stressors and perturbations.

## Concluding Remarks

>>>>

---
title: "Manuscript"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: 
  - Justin Pomeranz:
      email: jfpomeranz@gmail.com
      institute: cmu
      correspondence: true
  - name: James R. Junker
    institute: [glrc, lumcon]
  - name: Jeff S. Wesner
    institute: usd
institute:
  - cmu: Colorado Mesa University, Grand Junction, CO, USA
  - glrc: Great Lakes Research Center, Michigan Technological University, Houghton, MI USA
  - lumcon: Louisiana Universities Marine Consortium, Chauvin, LA USA
  - usd: Dept. of Biology, University of South Dakota, Vermillion, SD, USA
output:
   pdf_document:
     toc: yes
     keep_tex: true
     pandoc_args:
     - '--lua-filter=scholarly-metadata.lua'
     - '--lua-filter=author-info-blocks.lua'
   word_document:
      reference_docx: working_docx_template.docx
      toc: no
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
   html_document: default
      pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
header-includes:
  - \usepackage{lineno}
  - \usepackage{amsmath}
  - \usepackage{indentfirst}
  - \linenumbers
  - \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
indent: true
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

@@

For each sample of $M$ and each of the 1000 replicates, we estimated the exponent $\lambda$ using MLE methods modified from the the `sizeSpectra` package (Edwards et al. 2020). In addition, we use two common binning methods to estimate the OLS slope parameter in log-log space. For the first binning method, we created 6 equal logarithmic bins covering the range of body sizes. The count in each bin was normalized by dividing by the bin width. This method has been used by (refs here, I think a lot of Woodward's group uses this...) Throughout the manuscript, the normalized equal logarithmic binning method will be referred to as ELBn. The second method was similar to ELBn, but bins of Log~2~ widths are used. The count in each bin is normalized in the same way. This method has been used by [@pomeranz2018; @mcgarvey2019] and is referred to as the Normalized Abundance Spectrum (NAS).

@@

For each simulation replicate, we estimated how the size spectra parameters varied across the hypothetical gradient. Simple OLS regression were conducted in the form $\lambda_{estimate} ~ \beta_0 + \beta_1 * X$. The distribution of the relationship coefficient, $\beta_1$, were plotted compared to the known relationship. Likewise, the mean $\beta_1$ estimate ($\pm$ 1 SD) for all simulations were also plotted.

@@

We re-analyze two data sets of benthic macroinvertebrate communities from stream habitats across two different gradients. In the first, quantitative macroinvertebrate samples were collected from streams across an acid mine drainage stress gradient. Details of the sample collection and processing can be found in [@pomeranz2018]. Briefly, all individuals from each sample were identified to the lowest practical taxonomic unit and body lengths were measured using image processing software from photos taken with a camera mounted to a dissecting microscope. Body mass was estimated using taxon-specific published length weight regressions.

The second data set was from the wadeable stream sites of the National Ecological Observatory Network [@NEON_Inverts2022]. NEON stream sites are located across a wide temperature gradient in the United States, from Puerto Rico to Alaska. Quantitative macroinvertebrate samples were collected using the most appropriate method based on the local habitat. All individuals were identified and had their body lengths measured, and body mass was estimated using published length weight regressions. This data has been analyzed previously using size spectra methods as described in [@pomeranz2022]. Detailed methods of the sampling collection and data processing methods can be found on the NEON website ([https://www.neonscience.org/data-collection/macroinvertebrates], or macroinvertebrate DPI pubs).

@@

All methods performed reasonably well in detecting the known relationship across the hypothetical environmental gradient (mean $\beta_1$ estimates ranged from -0.419 to -0.498). However, the SD for the NAS and ELBn methods were 2-3x larger that that of the MLE method (Figure ~XX~). This result was consistent across simulation parameters, including sample size, number of sites, and $\lambda$ values.

![Mean estimate of the relationship coefficient for main simulation (A), with ten sites (B), with small variation in lambda (C), and with no relationship.Error bars indicate 1 standard deviation.](figures/slopes_combined.png)

@@

All of the methods performed similarly when there was no relationship across the hypothetical gradient. The mean $\beta_1$ estimates for all three methods ranged from -0.0005 to 0.0002. However, once again, the SD for the NAS and ELBn methods were ~3x larger than the SD for the MLE method.

@@

The performance of all methods to detect small variation in the $\lambda$ parameter (between -1.9 to -2.1) across the environmental gradient was also investigated. The mean $\beta_1$ coefficient estimates ranged from -0.10 to -0.09. The mean estimate for the MLE method was -0.1, which exaclty matched the value for the hypothetical relationship. Likewise, the upper 97.5 percentile of the distribution was below 0, meaning that most of the estimates had the correct sign. In contrast, the upper 97.5 percentile for both the NAS and ELBn methods were greater than 0 at 0.027 and 0.052, respectively. (Figure 4).

![Distribution of the estimated relationship for each method. Here, $\lambda$ varied from -1.9 to -2.1, with a known relationaship of -0.1 across the gradient indicated by the dashed vertical line.](figures/PLB_sim_small_relationship_density.png)

@@

Once again, the SD for both of the binning methods were $\ge$ 2.8 times that of the MLE method.

@@

![Need to load data on githib and rerun to make figure match new format above.](figures/neon_relationship.png)

@@

Measuring parameters describing the decline in abundance with increasing body size in communities is being done with increasing frequency across ecology. Previous work has investigated the accuracy and inherent biases associated with different estimation methods. However, how these innaccuracies and biases compound across environmental gradients remains uncertain, making it difficult to detect variation in size spectra parameters across environmental gradients with confidence. Here, we sampled body sizes from known distributions and estimated their parameters across a hypothetical gradient in order to assess the relationship coefficients obtained. Likewise, we compared how the interpretation of previously published results could change depending on the methodology used to estimate the relationship.

Generally, the estimate from the MLE method was always closer to the known relationship, and always had the smallest variation around the estimate. One of the benefits of using MLE is that all data points are retained for use within the model.

Binning methods are easy to use and interpret, which most likely accounts for their wide use in ecological studies. However, aggregating individuals into logarthmic bins removes a large source of the variation within the data by collapsing body size variation into a single values within bins. For example, all individuals placed into a bin that ranges from 2-4 grams of mass are all treated as having a mass of 3 grams, the midpoint of that bin. Likewise, a single abundance value is taken for this bin, despite that fact that there is almost certainly variation in the abundance of individuals that weigh ~2 grams or ~4 grams.  By homogenizing the data in this way, it seems that binning methods will always produce noisier results than MLE from the binning process alone, since binning is akin to deleting information that the model could otherwise use.

@@

Alarmingly, the performance of the ELBn and NAS all methods declined dramatically when trying to detect smaller variation in $\lambda$ values. We simulated observations from distributions varying by an absolute value of 0.2 units. This range of variation is similar to that which has been reported in many seminal works on variation in size spectra parameters.

Despite the drop in performance with reduced variation in $\lambda$ values, when a significant relationship was observed, it was generally in the correct direction and of a similar magnitude. This suggests that previously reported significant changes in size spectra parameters across environmental gradients and in experimental manipulations are plausible, and the magnitude of the relative change is a reasonable estimate. Given that all of the data within a study is treated identically, the the over all change in size spectra parameters is likely reasonable. However, the biases and inconsistencies in relationship estimates presented here suggest that it would be difficult if not impossible to compare the relative changes across different published studies which use different methods. Publication of individual body size data with future studies of size spectra relationships would greatly aid in our ability to generalize changes to this fundamental aspect of community organization across spatiotemporal scales and in response to environmental stressors and perturbations.

## Concluding Remarks
Both of the binning methods can perform as well as MLE under certain circumstances (large sample size, shallow $\lambda$). However, it would be extremely difficult to know if this was the case beforehand. Furthermore, in order to confirm this result, MLE methods would likely need to be employed. With the publication of the sizespectra package [-@edwards2017], producing MLE estimates of size spectra parameters is a relatively easy task. Therefore, we recommend using this in all future studies of size spectra relationships.
